Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-08 paper.tex
added one more comment of Bianca on potential of theoretical study of refinement

The Corona-virus warning app
%has taken a long time to develop being published only on
%16th June 2020. It
was a quite costly project but this was mainly due to the management of
Telekom and SAP being in the driving seat. But the app has been designed with great
attention on privacy:  a distributed architecture \cite{} has been adopted after a long and
heated debate with supporters of a central architecture. The distributed architecture is
based on a very clever distributed application design whereby users phones are sending
highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
protocol. The app saves those IDs of people in close proximity. When at a later date an
infected person reports his infection to a central server, the unique root ID is published
and in the daily check all mobile phones connecting to the central server download
the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
regularly check whether their user has been exposed to an infected person and issue
a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

\subsubsection{Refinement}
An additional feature that has been integrated into the Isabelle Infrastructure
framework motivated by security engineering formal specifications for IoT healthcare
system is an extension of the formal specification process introducing
refinement of Kripke structures \cite{kam:19a,kam:20a}. It refines a system model based on a 
formal definition of a combination of trace refinement and structural 
refinement (or datatype refinement). The definition allows to prove property preservation results 
crucial for an iterative development process.
The refinements of the system specification  can be interleaved with attack 
analysis while security properties can be proved in Isabelle. In each iteration
security qualities are accumulated while continuously attack trees scrutinize
the design.

>>>>

The Corona-virus warning app
%has taken a long time to develop being published only on
%16th June 2020. It
was a quite costly project but this was mainly due to the management of
Telekom and SAP being in the driving seat. But the app has been designed with great
attention on privacy:  a distributed architecture \cite{} has been adopted after a long and
heated debate with supporters of a central architecture. The distributed architecture is
based on a very clever distributed application design whereby users' phones are sending
highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
protocol. The app saves those IDs of people in close proximity. When at a later date an
infected person reports his infection to a central server, the unique root ID is published
and in the daily check all mobile phones connecting to the central server download
the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
regularly check whether their user has been exposed to an infected person and issue
a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

\subsubsection{Refinement}
An additional feature that has been integrated into the Isabelle Infrastructure
framework motivated by security engineering formal specifications for IoT healthcare
system is an extension of the formal specification process introducing
refinement of Kripke structures \cite{kam:19a,kam:20a}. It refines a system model based on a 
formal definition of a combination of trace refinement and structural 
refinement (or datatype refinement). The definition allows to prove property preservation results 
crucial for an iterative development process.
The refinements of the system specification  can be interleaved with attack 
analysis while security properties can be proved in Isabelle. In each iteration
security qualities are accumulated while continuously attack trees scrutinize
the design.
One of the contributions of this paper is to explore different concepts of refinement:
the formal expression of refinemnet, enables to pin down (i.e. exemplify) different concepts
of refinement (data refinement, action refinement, trace refinement (aka spec refinement) and
combinations thereof with concrete attack scenarios.
